Exploring the Validity Range of PLL Analytical Models

### 2025-06-02 CaseStudies/Example3_2_LimitsoftheTheorySoFar/Report/article/article.qmd
Nomeando corretamente as equações e gráficos.

This report presents the development of an Integer-N Phase-Frequency Detector/Charge Pump-based Phase-Locked Loop (PFD/CP-PLL) with a second-order loop filter. The loop filter is designed so that the loop bandwidth has a damping factor $\zeta=$ `{python} zeta`, and a 3-dB bandwidth of $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz. The example analyzed in this report is based on the case presented in @rogers2006IntegratedCircuitDesign.

@@

A [Fig. @fig-PLL-integerN-blocks] o diagrama de blocos de um PLL Integer-N PFD/CP-based.

@@

The loop filter transfer function, $F(s)$ can be described as:

$$
F(s) = \left(R+\frac{1}{sC_1}\right)\parallelsum\frac{1}{sC_2}
$$

@@

$$
G(s) = \frac{KF(s)}{s}
$$

@@

tal que $A_0$ é o ganho DC (usualmente unitário), $K_\theta=I/2\pi$ o ganho do charge pump, $K_\mathrm{VCO}$ o ganho do VCO e $N$ o divisor de frequência.

A função de transferência em malha fechada é dada por

$$
H(s)=\frac{\theta_\mathrm{fb}}{\theta_R}=\frac{\omega_\mathrm{fb}}{\omega_R}=\frac{G(s)}{1+G(s)}
$$ {#eq-PLL-closed-loop-transfer-function-initial}

@@

$$
E(s)=\frac{\theta_e}{\theta_R}=H(s)\cdot\frac{s}{KF(s)}
$$ {#eq-phase-error-transfer-function-initial}

@@

Analysing the maximum of the [Eq. @eq-inverse-laplace-phase-error], the time instant at which the maximum phase error occurs, $t_m$, can be computed using:

@@

[Eq. @eq-instant-maximum-frequency-step] represents the instant at which [Eq. @eq-inverse-laplace-phase-error] reaches its maximum value.

Sendo a máxima normalized phase error,

@@

$$
V(s)=\frac{V_C}{\theta_R}=H(s)\cdot\frac{1}{1/N}\cdot\frac{1}{K_\mathrm{VCO}}
$$ {#eq-voltage-control-transfer-function-initial}

@@

Calculando-se o instante em que ocorre a máxima diferença de fase obtém-se, $t_m=$ `{python} format(t_m)`s.

@@

Normalmente, espera-se que o máximo erro de fase suportado pelo PFD é de $2\pi$, entretanto, para coincidir com os valores obtidos com o simulador utilizado, empiricamente obtém-se um maximum phase error, $\theta_{e_m}=$ `{python} format_with_pi_latex(theta_e_m)`.

@@

A transfer function of the filter is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$F(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-filter-loop-transfer-function}}"))

@@

The open-loop transfer function of the system is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$G(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-open-loop-num-transfer-function}}"))
```

The closed-loop transfer function of the entire system, both in frequency domain and phase domain is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$H(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-closed-loop-num-transfer-function}}"))
```

The transfer function of the output phase is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$O(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-output-phase-num-transfer-function}}"))
```

The transfer function of the phase error is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$E(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-phase-error-num-transfer-function}}"))
```

The transfer function of the voltage control is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$V_C(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-voltage-control-num-transfer-function}}"))
```

@@

```{python}
#| echo: false
#| label: fig-pll-step-response
#| fig-cap: "Step Response of the Closed-Loop Frequency Transfer Function of the PLL."
#| fig-align: center

@@

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = 0.5e6
u = np.ones_like(t) * step_amplitude

@@

A step with the same amplitude as the previously calculated frequency step was also applied in phase, that is, $\Delta\theta=$ `{python} format_with_pi_latex(step_m_ang_meas)`rad/s. Applying this signal to [Eq. @eq-transfer-function], the response of this system to the phase step is obtained. With this step, the settling time achieved is $t_s=$ `{python} format(t_s_meas)`s. Through simulation, it is possible to verify that the obtained value is $t_s=$ `{python} format(ts_graphical)`s. The simulated response can be visualized in [Fig. @fig-pll-step-response].

```{python}
#| echo: false
#| label: fig-pll-step-response-alt
#| fig-cap: "Closed-loop phase step response of the designed PLL."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-pll-step-iasdhf-alt
#| fig-cap: "Closed-loop kjas step response of the designed PLL."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-pll-step-response-alt-double-ramp
#| fig-cap: "Closed-loop phase step response of the designed PLL."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-pll-step-response-1
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-phase-error-step-response
#| fig-cap: "Phase error response of step frequency."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-pll-step-response-2
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-pll-step-response-3
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-vc-step-response
#| fig-cap: "Voltage control step response of the designed PLL."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-pll-step-response-4
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-pll-step-response-5
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-pll-step-response-6
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

>>>>

This report presents the development of an Integer-N Phase-Frequency Detector/Charge Pump-based Phase-Locked Loop (PFD/CP-based PLL) with a second-order loop filter. The loop filter is designed so that the loop bandwidth has a damping factor $\zeta=$ `{python} zeta`, and a 3-dB bandwidth of $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz. The example analyzed in this report is based on the case presented in @rogers2006IntegratedCircuitDesign.

@@

A [Fig. @fig-PLL-integerN-blocks] apresenta o diagrama de blocos de um PLL Integer-N PFD/CP-based.

@@

The second-order loop filter transfer function, $F(s)$ can be described as:

$$
F(s) = \left(R+\frac{1}{sC_1}\right)\parallelsum\frac{1}{sC_2}
$$ {#eq-tf-filter}

@@

$$
G(s) = \frac{KF(s)}{s}
$$ {#eq-tf-open-loop}

@@

tal que $A_0$ é o ganho DC (usualmente unitário), $K_\theta=I/2\pi$ o ganho do charge pump, $K_\mathrm{VCO}$ o ganho do VCO e $N$ o divisor de frequência. A partir daí, é possível obter a função de transferência em malha fechada do sistema, dada por

$$
H(s)=\frac{\theta_\mathrm{fb}}{\theta_R}=\frac{\omega_\mathrm{fb}}{\omega_R}=\frac{G(s)}{1+G(s)}
$$ {#eq-tf-closed-loop}

É possível observar que a função de transferência $H(s)$ apresentada na [Eq. @eq-PLL-closed-loop-transfer-function-initial] vale tanto para a fase do sinal de feedback pela fase do sinal de referência quanto pela frequência do sinal de feedback pela frequência do sinal de referência.

@@

$$
E(s)=\frac{\theta_e}{\theta_R}=H(s)\cdot\frac{s}{KF(s)}
$$ {#eq-tf-phase-error}

@@

Neste momento, o interesse é obter uma equação que forneça o maior step de frequência que o PLL pode lidar. Para isso, analisa-se a equação do erro de fase e buscamos pelo valor de pico do erro de fase. Analysing the maximum of the [Eq. @eq-inverse-laplace-phase-error], the time instant at which the maximum phase error occurs, $t_m$, can be computed using:

@@

[Eq. @eq-instant-maximum-frequency-step] represents the instant at which [Eq. @eq-inverse-laplace-phase-error] reaches its maximum value. Sendo a máxima normalized phase error,

@@

$$
V_C(s)=\frac{V_C}{\theta_R}=H(s)\cdot\frac{1}{1/N}\cdot\frac{1}{K_\mathrm{VCO}}
$$ {#eq-tf-vctrl-vco}

## Output phase Transfer Function

A função de transferência da fase de saída pela fase de referência pode ser obtido como

$$
O(s)=\frac{V_C}{\theta_R}=H(s)\cdot\frac{1}{1/N}
$$ {#eq-tf-output-phase}

@@

Com os parâmetros $\zeta$ e $\omega_n$ definidos, é possível calcular o instante em que ocorre a máxima diferença de fase. Obtém-se, assim, $t_m=$ `{python} format(t_m)`s.

@@

Normalmente, espera-se que o máximo erro de fase suportado pelo PFD seja de $2\pi$. Entretanto, para coincidir com os valores obtidos com o simulador utilizado, empiricamente obtém-se um maximum phase error, $\theta_{e_m}=$ `{python} format_with_pi_latex(theta_e_m)`.

@@

Utilizando a [Eq. @eq-tf-filter], é possível obter the transfer function of the filter is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$F(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-filter-numeric}}"))

@@

Utilizando a [Eq. @eq-tf-open-loop], é possível obter the open-loop transfer function of the system is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$G(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-open-loop-numeric}}"))
```

Utilizando a [Eq. @eq-tf-closed-loop], é possível obter the closed-loop transfer function of the entire system, both in frequency domain and phase domain is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$H(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-closed-loop-numeric}}"))
```

Utilizando a [Eq. @eq-tf-output-phase], é possível obter the transfer function of the output phase is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$O(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-output-phase-numeric}}"))
```

Utilizando a [Eq. @eq-tf-phase-error], é possível obter the transfer function of the phase error is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$E(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-phase-error-numeric}}"))
```

Utilizando a [Eq. @eq-tf-vctrl-vco], é possível obter the transfer function of the voltage control is given by:

@@

# Exibe como equação no Jupyter
display(Markdown(f"$$V_C(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-vctrl-vco-numeric}}"))
```

@@

sim_stop_time = 50e-6
sim_points = 10000
step_amplitude = f_r
```

A [Fig. @fig-pll-step-response] apresenta um step em frequência de `{python} format(step_amplitude)`Hz aplicado a função de transferência da [Eq. @eq-tf-closed-loop]

```{python}
#| echo: false
#| label: fig-step-freq-h
#| fig-cap: "Frequency step response of the closed-loop transfer function."
#| fig-align: center

@@

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
u = np.ones_like(t) * step_amplitude

@@

A step with the same amplitude as the previously calculated frequency step was applied in [Fig. @fig-pll-step-response].

that is, $\Delta\theta=$ `{python} format_with_pi_latex(step_m_ang_meas)`rad/s. Applying this signal to [Eq. @eq-transfer-function], the response of this system to the phase step is obtained. With this step, the settling time achieved is $t_s=$ `{python} format(t_s_meas)`s. Through simulation, it is possible to verify that the obtained value is $t_s=$ `{python} format(ts_graphical)`s. The simulated response can be visualized in [Fig. @fig-pll-step-response].

```{python}
#| echo: false
#| label: fig-ramp-phase-h
#| fig-cap: "Phase response to a ramp input in the closed-loop transfer function."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-ramp-phase-h-o
#| fig-cap: "Comparison of phase responses to a ramp input in two different functions."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-ramp-phase-step-freq-h
#| fig-cap: "Phase response to a ramp with an abrupt frequency change (frequency step)."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-ramp-phase-h-phase-error
#| fig-cap: "Response to a ramp input with phase error."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-step-freq-e
#| fig-cap: "Phase error response to a frequency step input."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-step-freq-e-norm
#| fig-cap: "Normalized phase error response to a frequency step input."
#| fig-align: center

@@

@@

```{python}
#| echo: false
#| label: fig-step-freq-vc
#| fig-cap: "Control voltage response to a frequency step input."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-step-freq-vc-norm
#| fig-cap: "Normalized control voltage response to a frequency step input."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-bode-h
#| fig-cap: "Bode plot of the closed-loop transfer function."
#| fig-align: center

@@

```{python}
#| echo: false
#| label: fig-bode-f
#| fig-cap: "Bode plot of the filter transfer function."
#| fig-align: center
